---
title: "Environmental Influences on White-tailed Deer Cortisol Levels"
author: "Bryan Spencer"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

Read in necessary packages

```{r message=FALSE}
#install.packages(c('tidyverse', 'lme4', 'lmerTest', 'AICcmodavg', 'car', 'mgcv', 'MuMIn'))


library(tidyverse) #help with workflow
library(lme4) #mixed effect models
library(lmerTest) #mixed effect models with p-values
library(AICcmodavg) #rank models
library(car) #check model errors
library(MuMIn) #dredging to try ever combination of model parameters
library(rmarkdown)
library(here)
```

This chunk will read in dataset

```{r}

All <- read.csv('../raw/FinalFormat_EnvirCort_Females.csv', header = T) #Read in full dataset with Fecal Cort Levels, Environmental Data, and animal biological data

for(i in 1:nrow(All)){
  
  if(All$Age[i] == .5) {All$ageclass[i] <- "fawn"} else if(All$Age[i] %in% c(1.5, 2.5, 3.5)){All$ageclass[i] <- "imm"} else if(All$Age[i] %in% c(4.5, 5.5, 6.5))
    {All$ageclass[i] <- "mat"} #else {All$ageclass[i] <- "Mat"}
}


All <- All %>% mutate_at(19:29, scale) %>% #Scale and center all predictor variables
            filter(!is.na(Lactation)) %>% #remove any missing entries for lactation
            mutate_at(c("LeftMaster", "Year", "ageclass", "Capt_Loc", "Lactation"), factor) #Factor some of the biological data and random effects

All$Log_Fecal <- log(All$Fecal) #Log transform the fecal cort levels to approxiametly normality in the model errors




Does <- All %>% filter(Age %in% c(2.5, 3.5, 4.5, 5.5, 6.5)) #filter dataset to just sexually mature does 


#As the rut approaches, increase testosterone may influence buck cort levels. 
for(i in 1:nrow(Does)) { #For loop to cycle through each individual row
  
  #i <- 1
  Does$StudyJulian[i] <- as.Date(Does$Date[i]) - as.Date(paste0(Does$Year[i], '-10-01')) #Estimate the number of days since October 1 of that year
}

head(All)
#rain - milimeters
#biomass - kg/ha
#sand - %
#brush cover - %
#lactation - na(males), yes, no
#Gross BC (Boone & Crockett) - cm
#AdjWt - kg

```

This chunk cycles through rainfall periods to determine if timing of rain influence cort levels, and creates every possible model combination of specified parameters to determine the best fitting model for the data

```{r warning=F, message=F}

#this creates a vector of the column names containing important rainperiods
rain.var <- c("rain_1month_Previous", "rain_2month_Previous", "rain_3month_Previous", "rain_July_August")

#create empty list to store competitive models
f.model <- list()

#part of a progress bar
k <- 1

#This for loop cycles through each rain period variable and substitutes it the model formula
for (i in rain.var) {
  
  #paste the progress bar message
  cat('Working on', i, ':', k, 'of 6', '\n')
  
  #i <- rain.var[1] #debugger variable
  
  #the global model formula for the full data set and only sexually mature females
  f <- paste0('Log_Fecal ~ ageclass + Lactation*(', i, ' + Percent_Sand + WoodyCover + StudyJulian) + (1|Capt_Loc) + (1|Year)')
  
  #fit the global model with sexually mature females
  female <- lmer(data = Does, formula = as.formula(f), na.action = na.fail)

  #This creates every model for every parameter combination in the global model while specifying Ageclass as being in every model
  f.combo <- dredge(female)
  
  #This gets the top performing models after dredging with a delta AIC less than 5
  out3 <- get.models(f.combo, subset = delta < 5)
  
  for (x in 1:length(out3)){
    names(out3)[x] <- paste((as.character(formula(out3[[x]]))[3]), collapse = ' ')
  }
  
  #this saves outs to final list of top performing models
  f.model[[i]] <- out3
  
  #update the progress message
  k <- k + 1
}

f.model <- unlist(f.model)

test3 <- separate(data = data.frame(name = names(f.model)), col = name, into = c('var', 'model'), sep = '\\.')
test3$Dup <- !(duplicated(test3$model))

f.model <- f.model[test3$Dup]
names(f.model) <- test3$model[test3$Dup]
```
Evaluation of these models for sexually mature females shows that July-August rainfall is important influence on fall cort levels but the effect is dependent on lactation

```{r}

#for the sexually mature females the model with an interaction between lactation status and July-Aug rainfall was the best fitting
aic<-aictab(f.model)
write.csv(aic,'aicnew.csv')
#model summary for this top performing model
summary(f.model[['Lactation + rain_July_August + (1 | Capt_Loc) + (1 | Year) + Lactation:rain_July_August']])
summary(f.model[['WoodyCover + (1 | Capt_Loc) + (1 | Year)']])

```

This Chunk tells you how much cort changes (%) per cm increase in rain
```{r}

#lets get our coef from model summary
ef <- fixef(f.model[['Lactation + rain_July_August + (1 | Capt_Loc) + (1 | Year) + Lactation:rain_July_August']])

#Next calculate the slope of rain effect on cort of lactating female
slope <- ef["rain_July_August"] + ef["LactationYes - Lactating:rain_July_August"]

#This coef for rain is for a change in 1 SD of rain so we need to know this sd
ch.rain <- (1* attr(All$rain_July_August, 'scaled:scale') + attr(All$rain_July_August, 'scaled:center')) - (0* attr(All$rain_July_August, 'scaled:scale') + attr(All$rain_July_August, 'scaled:center'))

#so log(cort) is reduced -0.3743819 for every 52.24 mm increase in rain, now we want this change per cm of rain
crt.rain <- (slope)/(ch.rain/10)

#now to get an interpretable change in value in normal cort units 
final <- (exp(crt.rain) - 1)*100

#so this how much in % of how cort changes per cm increase in rain.
final

```


This chunk plots the effect above to visualize the impact

```{r}

library(ciTools) #needed to confidence intervals
library(colorBlindness)
library(ggplot2)

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

#creating new data to predict the effects of these models
df <- data.frame(matrix(nrow = 200, ncol = 2))
names(df) <- c('Lactation', 'rain_July_August')

df$rain_July_August <- seq(from = min(Does$rain_July_August), to = max(Does$rain_July_August), length.out = 100)
df$Lactation <- rep(unique(Does$Lactation), each = 100)

head(df)

#predict what the log fecal levels and the associated confidence intervals
df <- add_ci(df, f.model[['Lactation + rain_July_August + (1 | Capt_Loc) + (1 | Year) + Lactation:rain_July_August']], nSims = 15000, includeRanef = F)

df$rain_July_August <- df$rain_July_August * attr(All$rain_July_August, 'scaled:scale') + attr(All$rain_July_August, 'scaled:center') #unscale rainfall data
df <- df %>% mutate_at(vars(3:5), exp) #un-log transform the predicted cort levels

Does$rain_July_August <- Does$rain_July_August * attr(All$rain_July_August, 'scaled:scale') + attr(All$rain_July_August, 'scaled:center')

#plot
plot <- ggplot() +
  geom_ribbon(aes(x = df$rain_July_August, ymin = df$LCB0.025, ymax = df$UCB0.975, fill = df$Lactation), alpha = .35) +
  geom_line(aes(x = df$rain_July_August, y = df$pred, color = df$Lactation), lwd = 1.15) + 
  geom_point(aes(x = Does$rain_July_August, y = Does$Fecal, color = Does$Lactation)) + 
  ylab('FGM Concentrations (ng/g)') + 
  xlab('Rainfall (mm)') + 
  scale_color_manual(values = cbPalette[c(7,6)], name = 'Lactating', labels = c('No', 'Yes')) +
  scale_fill_manual(values = cbPalette[c(7,6)], name = 'Lactating', labels = c('No', 'Yes')) +
  theme_classic() + 
  theme(legend.position = c(.75, .75),
        axis.title = element_text(face = 'bold', size = 10),
        legend.title = element_text(face = 'bold', size = 8),
        legend.text = element_text(size = 7),
        axis.title.y = element_text(vjust = 1.5)
        )

colorbl <- cvdPlot(plot = plot)

ggsave(plot = colorbl, filename = 'CortPlot_ColBld.jpeg', path = 'G:/RProjects/Cort_East/Cort_SideProj/Plots', device = 'jpeg', dpi = 600, width = 7, height = 7, units = 'in')

ggsave(plot = plot, filename = 'CortPlot.jpeg', path = 'G:/RProjects/Cort_East/Cort_SideProj/Plots', device = 'jpeg', dpi = 600, width = 3.5, height = 3.5, units = 'in')

plot
```

